
## Read WHO Data

```{r}
WHO_DATA <- read.csv("C:/data/WHO.csv", stringsAsFactors = FALSE)
```

## Tidy Task 1:

using tidyr function gather()

```{r, echo=TRUE}

task1 <- gather(WHO_DATA,'new_sp_m014':'new_rel_f65',key="code",value="value" )
task1
dim(task1)

```

## Tidy Task 2:

```{r}

task2 <- task1%>%separate(code,into=c("new","var","sex_age"),sep="_")
task2.1 <- task2%>%separate(sex_age,into=c("sex","age"),sep=1)
task2.1
dim(task2.1)

```

## Tidy Task 3:

```{r}

task3 <- task2.1%>%spread(key="var",value="value")
task3
dim(task3)

```

## Tidy Task 4: 


```{r}

task4 <- task3%>%mutate(sex=factor(sex,levels=c("m","f")),age=factor(age,levels=c("014","1524","2534","3544","4554","5564","65"),labels  =c("<15","15-24","25-34","35-44","45-54","55-64", "65>="), ordered = T))
task4
dim(task4)

```

## Task 5: Filter & Select

```{r}

task5 <- task4 %>% select(., -c(iso2,new))
WHO_subset <- task5 %>% filter(country == "Albania"|country =="Algeria"|country =="Canada")
WHO_subset
dim(WHO_subset)

```

## Read Species and Surveys data sets

```{r}

species <- read.csv("C:/data/species.csv", stringsAsFactors = FALSE)
surveys <- read.csv("C:/data/surveys.csv", stringsAsFactors = FALSE)

```

## Task 6: Join  

In this task,left_join is used to avoid loss of data as inner_join will remove Specied_id(56 observations) with NA. This will result in data loss from other columns as well.

```{r}

surveys_combined <- left_join(surveys,species,by="species_id")
surveys_combined
dim(surveys_combined)

```

## Task 7: Calculate 
calculating the average weight and hindfoot length by choosing one of the species observed in each month

```{r}

task7 <- surveys_combined %>% filter(.,species_id == "PF")

task7 %>% group_by(month) %>% summarise(avgweight = mean(weight, na.rm=TRUE),avghindfoot = mean(hindfoot_length, na.rm=TRUE))


```

## Task 8: Missing Values

```{r}

surveys_combined_year <- surveys_combined %>% filter(., year == "1978")
surveys_combined_year %>% group_by(.,species) %>% summarise (total_na = sum(is.na(weight)))
surveys_weight_imputed<- surveys_combined_year %>% group_by(species) %>% mutate(weight =weight%>% impute(fun=mean))

surveys_weight_imputed %>% group_by(.,species) %>% summarise (total_na = sum(is.na(weight)))

surveys_weight_imputed %>% summarise (totalinf = sum(is.infinite(weight)))
surveys_weight_imputed %>% summarise (totalNaN = sum(is.nan(weight)))

#Checking outliers using tukey's method
boxplot(surveys_combined$hindfoot_length,main="Box Plot of Hindfoot length", ylab="hindfoot length", col = "grey")
dim(surveys_combined)
#Getting actual values of outliers
boxplot(surveys_combined$hindfoot_length)$out

#assign the outliers into vector
outliers_Vector <- boxplot(surveys_combined$hindfoot_length)$out
print(outliers_Vector)

# finding the rows in which outliers are present
surveys_combined[which(surveys_combined$hindfoot_length %in% outliers_Vector),]

# Removing outliers 
surveys_combined <- surveys_combined[-which(surveys_combined$hindfoot_length %in% outliers_Vector),]

#checking results
boxplot(surveys_combined$hindfoot_length,main="Box Plot of Hindfoot length", ylab="hindfoot length", col="red")
dim(surveys_combined)


